---
title: 'Statistical Inference: Part 1- Comparing the exponential distribution in R and the Central Limit Theorem '
author: 'Lawrence Lau'
output:
  html_document:
      keep_md: yes
---

This is Part 1 of the peer assessment project from the Statistical Inference class in the Data Scientist Specialization course.  The objective is to explore the exponential distribution in R and compare it with the Central Limit Theorem.  

### <font color=blue> Simulations </font>

Below, we generate a histogram of 40 exponentials using rate(lambda) of 0.2.  The theoretical mean and standard deviation of an exponential distribution are 1/lambda, so we expect to see both somewhere around 5.  
```{r}
a <- rexp(40, rate=.2)
ma <- mean(a)
sda <- sd(a)
hist(a, main="Histogram of 40 exponentials")
abline(v=ma, col="blue", lwd=2)
```

We see the **sample mean**, denoted by the vertical blue line, is **`r ma`**. Remember, the theoretical mean and standard deviation of an exponential distribution are 1/lambda (1/.2 = 5). In this sample mean, we're `r 5-ma` off. The **sample standard deviation** is **`r sda`**, which is off by `r 5-sda`.  

Below is a histogram of the distribution of averages of 40 exponentials performed 1,000 times. Since it's taking 1,000 averages of 40 randomly drawn exponentials (normalizing the distribution), we expect to see this histogram's shape take a bell-curve shape.  We also expect the mean to be closer to the 1/lambda calculation of the theoretical mean. and standard deviation-  certainly much closer than the previous histogram where it was just 1,000 randomly drawn exponentials.  
```{r}
mns=NULL
for (i in 1:1000) mns=c(mns, mean(rexp(40, rate=.2)))
hist(mns, main="Histogram of 1,000 Averages of 40 exponentials")
maa <- mean(mns)
sdaa <- sd(mns)
abline(v=maa, col="blue", lwd=2)
```

We see here the gaussian (bell-curved) nature of the histogram, which is a criterion of a normal distribution.  Indeed, the Central Limit Theorem states that "the mean of any set of variates with any distribution having a finite mean and variance tends to the normal distribution".* We also see the mean, once again denoted by a vertical blue line, is `r maa`. This confirms the theoretical mean of 1/lambda.  The standard deviation is `r sdaa`, which also confirms the theoretical standard deviation.   



*http://math.about.com/od/glossaryofterms/g/Bell-Curve-Normal-Distribution-Defined.htm
